Notes on the Godunov Step in {\tt MAESTRO} and {\tt VARDEN}

### 2008-04-18 docs/Godunov_notes/paper.tex
more notes about VARDEN and MAESTRO differences

These are working notes for the Godunov step in {\tt MAESTRO} and {\tt VARDEN}.
\section{Differences Between These Notes and the Code}
\begin{enumerate}
\item Concerning both equations (\ref{Radial utrans_L}) and (\ref{Radial utrans_R}), in the code we are omitting the terms in equation (\ref{Perturbational Velocity Equation}) proportional to $w_0$ (by using $\wt$ rather than $w$ as the advection speed) and $(\Ubt\cdot\eb_r)$ (by omitting the term altogether).
\item Concerning (\ref{Force Term}), in the code we are not including the term proportional to $\partial\pi_0/\partial r$.
\end{enumerate}
\section{Questions}
\begin{enumerate}
\item We need to re-examine when to include the term proportional to $(\Ubt\cdot\eb_r)$ in equation (\ref{Perturbational Velocity Equation}) when we do a 1D extrapolation of the radial velocity to r-faces.  For example, equations (\ref{Radial utrans_L}) and (\ref{Radial utrans_R}) include it, but equations (\ref{1D wt_L extrap}) and (\ref{1D wt_R extrap}) do not.
\item In equations (\ref{wt_L Extrapolation}) and (\ref{wt_R Extrapolation}), (\ref{wt_L x-edge}) and (\ref{wt_R x-edge}), (\ref{wt_L r-edge}) and (\ref{wt_R r-edge}), we could also compute the term proportional to $\partial w_0/\partial r$ by using the cell-centered $\wt$ rather than the average of the neighboring $\wt^{\trans}$.
\end{enumerate}
\section{Differences Between {\tt MAESTRO} and {\tt VARDEN}}
\begin{enumerate}
\item {\tt VARDEN} does not use the concept of $w_0$, so there is no such thing as $\Ubt$ either.  {\tt VARDEN} only does Cartesian problems.
\item {\tt VARDEN} does not use the concept of $\Ub^{\trans}$, but rather in {\tt advance\_premac}, the function {\tt velpred} computes $\Ub^{\mac,*}$ directly.
\item {\tt VARDEN} uses characteristic tracing in the Taylor series extrapolation for $\Ub^{\mac,*}$.
\item {\tt VARDEN} uses an ``FPU'' type algorithm to get both $\Ub^{\edge}$ and $s^{\edge}$, whereas {\tt MAESTRO} uses the ``FPU'' algorithm for $s^{\edge}$, but not $\Ub^{\edge}$
\item The transverse derivative terms for $\rho$ in {\tt VARDEN} are done in a conservative formulation, rather than advective.
\end{enumerate}
\cleardoublepage
\section{Computing $\Ubt^{\trans}$ in {\tt MAESTRO}}
In {\tt advance\_premac}, we call the function {\tt mkutrans}, which computes the edge-centered transverse velocities, $\Ubt^{\trans}$.  These transverse velocities do not contain $w_0$, so immediately following the call to {\tt mkutrans}, we call {\tt addw0} to compute $\Ub^{\trans}$ from $\Ubt^{\trans}$.  We will only compute the normal component of velocity at each face.\\

@@

\section{Computing $\rho^{'\edge}, X_k^{\edge}, h^{\edge},$ and $T^{\edge}$ in {\tt MAESTRO}}
In {\tt scalar\_advance}, we call {\tt make\_edge\_scal} to compute $\rho^{'\edge}, X_k^{\edge}, h^{\edge},$ and $T^{\edge}$ at each edge.  The procedure is the same for each scalar, so we shall simply denote the scalar as $s$.  We always need to compute $\rho'$ and $X_k$ to edges.

@@

f_T &=& \frac{1}{\rho c_p}\left\{(1-\rho h_p)\left[\psi + (\Ubt\cdot\eb_r)\frac{\partial p_0}{\partial r}\right] - \sum_k\rho\xi_k\dot\omega_k + \rho H_{\rm nuc} + \rho H_{\rm ext} + \nabla\cdot\kth\nabla T \right\}.\nonumber \\

@@

u_{L,\ib-\half\eb_x}^{\mac,*} &=& u_{L,\ib-\half\eb_x}^{1D} - \frac{\dt}{4h}\left(v_{\ib-\eb_x+\half\eb_y}^{1D}+v_{\ib-\eb_x-\half\eb_y}^{1D}\right)\left(u_{\ib-\eb_x+\half\eb_y}^{1D} - u_{\ib-\eb_x-\half\eb_y}^{1D}\right) + \frac{\dt}{2}f_{u,\ib-\eb_x}, \nonumber \\
&& \\

@@

s_{L,\ib-\half\eb_x}^{\edge} &=& s_{L,\ib-\half\eb_x}^{1D} - \frac{\dt}{4h}\left(v_{\ib-\eb_x+\half\eb_y}^{\mac}+v_{\ib-\eb_x-\half\eb_y}^{\mac}\right)\left(s_{\ib-\eb_x+\half\eb_y}^{1D} - s_{\ib-\eb_x-\half\eb_y}^{1D}\right) + \frac{\dt}{2}f_{s,\ib-\eb_x}, \nonumber \\
&& \\

>>>>

These are working notes for the Godunov step in {\tt MAESTRO} and {\tt VARDEN}.
\section{Differences Between These Notes and the Code}
\begin{enumerate}
\item Concerning both equations (\ref{Radial utrans_L}) and (\ref{Radial utrans_R}), in the code we are omitting the terms in equation (\ref{Perturbational Velocity Equation}) proportional to $w_0$ (by using $\wt$ rather than $w$ as the advection speed) and $(\Ubt\cdot\eb_r)$ (by omitting the term altogether).
\item Concerning (\ref{Force Term}), in the code we are not including the term proportional to $\partial\pi_0/\partial r$ in the edge-state prediction.  However, this term is present in the cell-update.
\end{enumerate}
\section{Questions}
\begin{enumerate}
\item We need to re-examine when to include the term proportional to $(\Ubt\cdot\eb_r)$ in equation (\ref{Perturbational Velocity Equation}) when we do a 1D extrapolation of the radial velocity to r-faces.  For example, equations (\ref{Radial utrans_L}) and (\ref{Radial utrans_R}) include it, but equations (\ref{1D wt_L extrap}) and (\ref{1D wt_R extrap}) do not.
\item In equations (\ref{wt_L Extrapolation}) and (\ref{wt_R Extrapolation}), (\ref{wt_L x-edge}) and (\ref{wt_R x-edge}), (\ref{wt_L r-edge}) and (\ref{wt_R r-edge}), we could also compute the term proportional to $\partial w_0/\partial r$ by using the cell-centered $\wt$ rather than the average of the neighboring $\wt^{\trans}$.
\end{enumerate}
\section{Differences Between {\tt MAESTRO} and {\tt VARDEN}}
\begin{enumerate}
\item {\tt VARDEN} does not use the concept of $w_0$, so $\Ubt$ does not exist.
\item {\tt VARDEN} only does Cartesian problems.
\item {\tt VARDEN} does not use the concept of $\Ub^{\trans}$, but rather in {\tt advance\_premac}, the function {\tt velpred} computes $\Ub^{\mac,*}$ directly.
\item {\tt VARDEN} uses characteristic tracing in the Taylor series extrapolation for $\Ub^{\mac,*}$, whereas {\tt MAESTRO} does not.
\item {\tt VARDEN} uses an ``FPU'' type algorithm to get both $\Ub^{\edge}$ and $s^{\edge}$, whereas {\tt MAESTRO} uses the ``FPU'' algorithm for $s^{\edge}$, and a ``non-FPU'' algorithm for $\Ub^{\edge}$
\item {\tt VARDEN} treats $\rho$ in a conservative manner, whereas {\tt MAESTRO} treats $\rho'$ in a non-conservative manner.  Therefore, {\tt MAESTRO} accounts for this difference by modifying the forcing term in {\tt modify\_scal\_force}
\item {\tt MAESTRO} does not corner couple for 3D problems.
\end{enumerate}
\cleardoublepage
\section{Computing $\Ubt^{\trans}$ in {\tt MAESTRO}}
In {\tt advance\_premac}, we call the function {\tt mkutrans}, which computes the edge-centered transverse velocities, $\Ubt^{\trans}$.  These transverse velocities do not contain $w_0$, so immediately following the call to {\tt mkutrans}, we call {\tt addw0} to compute $\Ub^{\trans}$ from $\Ubt^{\trans}$.  We will only compute the normal component of velocity at each face.\\

@@

\section{Computing $\rho^{'\edge}, X_k^{\edge}$, and $(\rho h)^{'\edge}/h^{\edge}/T^{\edge}$ in {\tt MAESTRO}}
In {\tt scalar\_advance}, we call {\tt make\_edge\_scal} to compute $\rho^{'\edge}, X_k^{\edge}$, and $(\rho h)^{'\edge}/h^{\edge}/T^{\edge}$ at each edge.  The procedure is the same for each scalar, so we shall simply denote the scalar as $s$.  We always need to compute $\rho'$ and $X_k$ to edges, and the choice of energy prediction is as follows:

@@

f_T &=& \frac{1}{\rho c_p}\left\{(1-\rho h_p)\left[\psi + (\Ubt\cdot\eb_r)\frac{\partial p_0}{\partial r}\right] - \sum_k\rho\xi_k\dot\omega_k + \rho H_{\rm nuc} + \rho H_{\rm ext} + \nabla\cdot\kth\nabla T \right\}.

@@

u_{L,\ib-\half\eb_x}^{\mac,*} &=& u_{L,\ib-\half\eb_x}^{1D} - \frac{\dt}{4h}\left(v_{\ib-\eb_x+\half\eb_y}^{1D}+v_{\ib-\eb_x-\half\eb_y}^{1D}\right)\left(u_{\ib-\eb_x+\half\eb_y}^{1D} - u_{\ib-\eb_x-\half\eb_y}^{1D}\right) + \frac{\dt}{2}f_{u,\ib-\eb_x}, \\

@@

s_{L,\ib-\half\eb_x}^{\edge} &=& s_{L,\ib-\half\eb_x}^{1D} - \frac{\dt}{4h}\left(v_{\ib-\eb_x+\half\eb_y}^{\mac}+v_{\ib-\eb_x-\half\eb_y}^{\mac}\right)\left(s_{\ib-\eb_x+\half\eb_y}^{1D} - s_{\ib-\eb_x-\half\eb_y}^{1D}\right) + \frac{\dt}{2}f_{s,\ib-\eb_x}, \\
